---
title: "Coursera Assignment"
author: "Rajesh More"
date: "March 6, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reproducible Research: Peer Assessment 1
**1. Loading and preprocessing the data**

Read raw data into input_raw_data varable
Load the data (i.e. ğš›ğšğšŠğš.ğšŒğšœğšŸ())
Process/transform the data (if necessary) into a format suitable for your analysis

```{r input_raw_data}
library(lattice)
setwd("~/Documents/Coursera/Reproducible Research/week1/assignment")
input_raw_data <- read.csv("activity.csv",header = TRUE,sep = ",")
input_raw_data$date <- as.Date(input_raw_data$date, "%Y-%m-%d")
input_raw_data_new <- input_raw_data
summary(input_raw_data)
```

**2. What is mean total number of steps taken per day?**

Calculate the total number of steps taken per day. Get total steps per day - using tapply (group by date and find sum for steps  )

Check bar plot - barplot(totstepsbydate)
Plot histogram

```
totstepsbydate <- tapply(input_raw_data$steps, input_raw_data$date, sum)
hist(totstepsbydate,  xlab = "Steps Per Day", ylab = "Interval", main = "Steps per day")
```
``` {r , echo=FALSE}
totstepsbydate <- tapply(input_raw_data$steps, input_raw_data$date, sum)
hist(totstepsbydate,  xlab = "Steps Per Day", ylab = "Interval", main = "Steps per day")
```

**Remove missing values (remove .NA)**

```{r , echo=FALSE}
totstepsbydate <- totstepsbydate[!is.na(totstepsbydate)]
hist(totstepsbydate,  xlab = "Steps Per Day", ylab = "Interval", main = "Steps per day")
```

Find mean and median
```{r,echo=TRUE}
meanvalue <- mean(totstepsbydate)
meanvalue
medianvalue <- median(totstepsbydate)
medianvalue
```

**3. What is the average daily activity pattern?**
```{r,echo=TRUE}
# To find average get mean 
meansteps <- tapply(input_raw_data$steps, input_raw_data$interval, mean,na.rm = TRUE)
# Plot line graph for Average daily activity pattern
plot(row.names(meansteps), meansteps, type = "l", xlab = "5 min interval", 
 ylab = "Average steps/day", main = "Average Daily Activity Pattern", col = "dark blue", lwd = 2 )
```

**4. Imputing missing values**

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with ğ™½ğ™°. s)Use complete.cases - it return a logical vector indicating which cases are complete, i.e., have no missing values.
Get NA set with !

```{r}
nrow(input_raw_data_new[!complete.cases(input_raw_data_new),])

for (i in 1:dim(input_raw_data_new)[1]) {
   ifelse (is.na(input_raw_data_new$steps[i]) ,   input_raw_data_new$steps[i] <- mean(meansteps) ,NA)
  }
```
Check record count with and without NA
```{r}
# rows with NA values
nrow(input_raw_data[!complete.cases(input_raw_data),])
# rows with NA values (NA removed)
nrow(input_raw_data_new[!complete.cases(input_raw_data_new),])
```
Histogram after removing NA
```{r,echo=TRUE}
totstepsbydate_new <- tapply(input_raw_data_new$steps, input_raw_data_new$date, sum)
hist(totstepsbydate_new,  xlab = "Steps Per Day", ylab = "Interval", main = "Steps per day")
```
Find mean and median value with and without NA
```{r}
meanvalue_new <- mean(totstepsbydate_new)
# Mean with NA
meanvalue
# Mean without NA
meanvalue_new
medianvalue_new <- median(totstepsbydate_new)
# Median with NA
medianvalue
# Median with NA
medianvalue_new
```

**5. Are there differences in activity patterns between weekdays and weekends?**

```{r}
day <- weekdays(input_raw_data_new$date)
week_weekend <- vector()
for (i in 1:nrow(input_raw_data_new)) {
  if (day[i] == "Saturday") {
    week_weekend[i] <- "Weekend"
  } else if (day[i] == "Sunday") {
    week_weekend[i] <- "Weekend"
  } else {
    week_weekend[i] <- "Weekday"
  }
}
```

```{r,echo=TRUE}
# Add daylevel to input_raw_data_new
input_raw_data_new$week_weekend <- week_weekend
input_raw_data_new$week_weekend <- factor(input_raw_data_new$week_weekend)

# split data and find mean steps for interval/day
stepsperday <- aggregate(steps ~ interval + week_weekend, data = input_raw_data_new, mean)
names(stepsperday) <- c("interval", "week_weekend", "steps")
# Plot week vs weekend chart
xyplot(steps ~ interval | week_weekend, stepsperday, type = "l", layout = c(1, 2),xlab = "interval", ylab = "steps")
```


